---
output: html_document
---
Reproducable Results Assignment 1

whutton
================================

Load and Process the Data

Histogram of the total number of steps taken each day
```{r}
library(dplyr)
bdata <- read.csv("c:\\hutton\\r_reproduce\\activity.csv")
zip2 <- group_by(bdata,date)
zip2a <- group_by(bdata,date)
zip3 <- summarise(zip2,step_cnt = sum(steps,na.rm = TRUE))
daily <- summarise(zip2a,step_cnt = sum(steps,na.rm = TRUE),
                   max1 = max(steps,na.rm=TRUE),
                   mean = mean(steps,na.rm=TRUE),
                   medium = median(steps[steps>0])
                     )
hist(daily$step_cnt,
     main="Total Number of Steps taken each day",
     xlab = "Daily Step Count")

```


Mean - medium steps taken each day

```{r}
#print(tbl_df(daily[,c("date","mean","medium")]),n=100)
```

Average daily activity

```{r}
t1 <- group_by(bdata,interval)
t2 <- summarise(t1,avg_period = mean(steps,na.rm=TRUE) )


plot(t2$interval,t2$avg_period,type="l",xlab="Interval",ylab="All Days")

```
Period with the max average steps
```{r}
max_average <- subset(t2,avg_period==max(t2$avg_period,na.rm=TRUE))
max_average
```

Number of Rows with Missing Values

```{r}
sum(!complete.cases(bdata))
```

Inputting Missing Values

```{r}
rows_in_data <- nrow(bdata)
bdata1 <- bdata
for ( i in 1:rows_in_data ) {
	step_value <- bdata1[i,"steps"]
	if (is.na(step_value)) {
		missing_interval <- bdata1[i,"interval"]
		value1 <- t2[which(t2$interval == missing_interval),2]
		bdata1[i,"steps"] = value1
      }
}


zip2a <- group_by(bdata1,date)
zip3 <- summarise(zip2a,step_cnt = sum(steps,na.rm = TRUE))
daily1 <- summarise(zip2a,step_cnt = sum(steps,na.rm = TRUE),
                   max1 = max(steps,na.rm=TRUE),
                   mean = mean(steps,na.rm=TRUE),
                   medium = median(steps[steps>0])
                     )
hist(daily1$step_cnt,
     main="Historgram Daily Steps with Missing Values Added",
     xlab="Daily Step Counnt")


```

Mean and median of the total number of steps taken per day

```{r}
print(tbl_df(daily1[,c("date","mean","medium")]),n=100)
```

Inputting missing values

```{r}
new_daily <- merge(daily,daily1,by ="date",all=TRUE)
daily1[] <- lapply(daily1,as.character)

```
 
```{r sum_missing}
sum_missing <- sum(new_daily$step_cnt.x,na.rm = TRUE)
```

```{r sum_avg}
sum_avg <- sum(new_daily$step_cnt.y,na.rm = TRUE)
```

Total number of steps (with missing values) `r sum_missing`

Total number of steps (with Average substitution) `r sum_avg`

Activiity broken out by Weekday / Weekend

```{r}
weekDay <- function(x) {
	date1 <- weekdays(as.Date(x))
	return_value <- "Weekday"
	if (date1 == "Sunday") {
		return_value <- "Weekend"
	}
	if (date1 == "Saturday") {
		return_value <- "Weekend"
	}
	return_value
}


zip22 <- group_by(bdata1,interval)


rows_in_data <- nrow(zip22)

for ( i in 1:rows_in_data ) {
	value1 <- weekDay(zip22[[i,"date"]])

	zip22[i,"dayType"] = value1
}

zipWeekDay <- subset(zip22,dayType=="Weekday")
zipWeekEnd <- subset(zip22,dayType=="Weekend")

daily10 <- summarise(zipWeekDay,step_cnt = sum(steps,na.rm = TRUE),
                   max1 = max(steps,na.rm=TRUE),
                   mean = mean(steps,na.rm=TRUE),
                   medium = median(steps[steps>0])
                     )
daily11 <- summarise(zipWeekEnd,step_cnt = sum(steps,na.rm = TRUE),
                   max1 = max(steps,na.rm=TRUE),
                   mean = mean(steps,na.rm=TRUE),
                   medium = median(steps[steps>0])
                     )

par(mfrow=c(2,1))
plot(daily10$interval,
     daily10$mean,
     type="l",xlab="Interval",ylab="Weekdays")
plot(daily11$interval,
     daily11$mean,
     type="l",xlab="Interval",ylab="Weekends")
```
